---
title: "Gaussian_Practice"
author: "Jon Nations"
date: "4/8/2020"
output: html_document
---

### The goal of this project is to figure out how to do better comparative work.
####The way to do that is:
1) Figure out what brms is doing with ```cov_ranef```, is it infact Brownian Motion?
2) See what the Stan code for a similar model is in Rethinking, but using the other Gaussian Process regression codes (Not Brownian Motion)
3) See if I can use my Stan script from a brms model and replace the ```cov_ranef``` with one of the other Gaussian Processes from Rethinking 2nd ed. Chapter 14.
#### Setup
```{r set root}
knitr::opts_knit$set(root.dir = "~/Documents/Projects/Covariance/Covariance")
```

```{r libraries}
library(tidyverse)
library(rethinking)
library(brms)
library(ape)
```

```{r data}
data("Primates301")
data("Primates301_nex")
```

#### Compare BRMS Stancode to Rethinking Stancode
I need to know what brms is doing. I think I will use the same models from Rethinking in brms so I don't have to try to switch the other way, from brms to rethinking. 

Get Data sorted
```{r data}
d <- Primates301
d$name <- as.character(d$name)
dstan <- d[ complete.cases(d$group_size, d$body, d$brain), ]
spp_obs <- dstan$name

# Rethinking data -----
dat_list <- list(N_spp = nrow(dstan),
                 M = standardize(log(dstan$body)),
                 B = standardize(log(dstan$brain)),
                 G = standardize(log(dstan$group_size)),
                  Imat = diag(nrow(dstan)))

# BRMS Data -----
dstan$M = standardize(log(dstan$body))
dstan$B = standardize(log(dstan$brain))
dstan$G = standardize(log(dstan$group_size))
```

Get tree sorted
Interestingly, Rethinking called the corBrownian command before getting the VCV matrix, however, I can't see a single difference between the matrices V and V2 below.
```{r tree}

tree_trimmed <- keep.tip(Primates301_nex, spp_obs)
#Brownian Tree
Rbm <- corBrownian(phy = tree_trimmed)
V <- vcv(Rbm)
dat_list$V <- V[spp_obs, spp_obs]
dat_list$R <- dat_list$V / max(V)
# Non-Brownian Tree, but same thing!
V2 <- vcv(tree_trimmed)
```

BRMS Code
```{r brms bm14.10}
bm14.10 <- brm(B ~ M + G + (1|name),
               data = dstan,
               family = gaussian(),
               cov_ranef = list(name = V2),
                 prior = c(
                           prior(normal(0, 1), "Intercept"),
                           prior(normal(0, 0.5), "b"),
                           prior(exponential(1), "sigma")),
              cores = 4)
```

```{r}
summary(bm14.10)
```

Rethinking Code
```{r rethinking rm14.10}
rm14.10 <- ulam(
  alist(
    B ~ multi_normal( mu, SIGMA),
    mu <- a + bM*M + bG*G,
    matrix[ N_spp, N_spp]: SIGMA <- R * sigma_sq,
    a ~ normal(0, 1),
    c(bM, bG) ~ normal(0, 0.5),
    sigma_sq ~ exponential(1)),
    data = dat_list, chains = 4, cores = 4)
```

```{r}
precis(rm14.10)
```

